Self-Supervised Learning of Generalized Policies in adversarial domains

### 2021-06-01 doc/paper/paper.tex
minor change in format
minor change
doc update

\subsection{Space Invaders}
\subsubsection{Description}
In this domain the agent can $shoot$, move to the $right$ or to the $left$. The agent during his turn can do two actions, while the Martians can do one. Note that if the Martians and the agent have only one action per turn the agent cannot win, this is because of the situation where the agent stands in the Martian's column to shoot him on his next turn, but the next turn (the Martian's turn) the Martian can move away, this leads to an infinite loop if both are playing perfectly. That is, the agent must be able to move to the agent's column and then shoot. To win the agent has to kill all the martians, for this he must shoot each of them. The agent kills an opponent if he shoots while being in the same column, and he can only kill one for each shoot. The martians has three possible actions $left$, $right$, $up$ and $down$. All the adversaries move in the same direction. The agent cannot be destroyed in any way by the Martians, they can go down to the row above the agent. The agent cannot collide with the Martians as they cannot descend to the row above it. Note that the policy found is optimal and will solve all possible instances of the game. The policy is also optimal for a version of the game where the Martians destroy the agent when they reach the row above him.

@@

\subsubsection{Policy Description}
According to the first rule if the agent is not in the column of a Martian, $B$ is greater than zero, it is good to decrement $B$, i.e. move towards the closest martian. The second rule is applicable when the agent is under a Martian, in the same column, and according to this rule it is good to decrease the number of Martians, i.e. execute the action $shoot$.

@@

\subsection{Nim}
\subsubsection{Description}
In this game the action space is equal to the number of tokens $t$ in the grid. In order to win the agent has to remove the last token in his turn. If the opponet removes the last token in his turn, the agent loses. The agent and the adversary has on action per turn. When one token is removed, the tokens above in the same column are removed too.
It is interesting to note that in this game not all states are solvable if the opponent plays optimally. Knowing that the expert policy is to leave the opponent with both stacks with the same number of tokens, this state with both stacks leveled will not be solvable. Therefore, in order to solve it, we have made use of the $Bad(state)$ feature, which allows us not to have to define a good transition for each state, and as an alternative to label it as a bad state. The advantage is that if the initial state is solvable, bad states can always be avoided. However the policy is not complete for all states, in particular there will be no good transition for states with the two token stacks leveled, as we will see below, which we can say is the optimal representation of the game strategy.

@@

\subsubsection{Features}
\begin{Verbatim}[fontsize=\footnotesize]
  (A) Num[And(cell-has-token,Exists(left_cell,cell-has-token))] [k=5]
  (B) Num[And(cell-has-token,Forall(leftdown_cell,cell-has-token))] [k=5]
  (C) Num[And(cell-has-token,Forall(rightdown_cell,cell-has-token))] [k=5]
\end{Verbatim}

@@

\subsubsection{Policy}
\begin{Verbatim}[fontsize=\footnotesize]
  1. B>0 AND C>0 -> {A ·, B ↓, C ↓}
\end{Verbatim}

\subsubsection{Policy Description}
The policy can be interpreted as leaving the two token stacks with the same number of tokens. This is done by decreasing the number of tokens that have a token to their leftdown or rightdown without decreasing the number of tokens with a token next to them.
As the policy clearly shows, there is no good transition, that includes a change in the number of tokens that a token has on its left. In other words, states where the same number of tokens in the two stacks cannot be preserved with some legal move are bad, and therefore they have no good transition. In fact according to the policy it is good to leave the opponent in these states bad, because there will be no legal move for the opponent to leave us with the same number of tokens in the two stacks.

@@

In this domain the agent has to pick up a missed pet and take it to a certain location, the destination. The pet has the option on its turn to escape from its captor to any of the cells adjacent to the pet, even if is holded by the agent it can escape. The goal is defined by the pet being droped at the destination cell, not holded by the agent. The agent has two actions per turn and the adversary one, otherwise the game would not be solvable for all the states. The set of possile actions for the agent is ~\{$pick$, $drop$, $rightup$, $reightdown$, $leftup$, $leftdown\}$. The agent, in order to pick up or drop the pet has to execute the actions $pick$ or $drop$ respectively being at the pet's cell. The set of possile actions for the pet consist in escape to one of the adjacent cells to it, i.e. at most eight possible actions depending on the position of the pet in the grid. The pet can not move into the delivery point.

@@

\subsubsection{Policy Description}
According to the first rule, it is good to move towards the pet if the agent is not holding it and the distance to it is greater than zero. According to the second rule, if the agent has the pet, and is at the destination, it is good to move towards the destination. According to the third rule of the policy, if the agent is holding the pet, and is at destination it is good to drop the pet. Finally, according to the fourth rule of the policy, if the agent does not have the pet, and is at the same cell it is good to pick it.

\subsubsection{Example of Game}
\begin{Verbatim}[fontsize=\footnotesize]
 # p.... # p.... # .p... # .p... # ..... # ..... # ..... # ..... # ..... # ..... # ..... # ..... # ..... #
 # ..... # ..... # A.... # ..... # ..... # ..... # ..... # ..A.. # .pA.. # ..A.. # ..... # ..... # ..... #
 # ..... # ..... # ..... # .A... # .A... # pA... # .A... # ..... # ..... # ..... # ..... # ..... # ..... #
 # ..... # .A... # ..... # ..... # ..... # ..... # ..... # ..... # ..... # ..... # ...A. # ..... # ..... #
 # A...D # ....D # ....D # ....D # ....D # ....D # ....D # ....D # ....D # ....D # ....D # ....A # ....A #
\end{Verbatim}

\subsection{Shoot}
\subsubsection{Description}
In this domain an agent has to shoot an adversary. The goal is defined by the agent being alone on the grid, having finished with the opponent. The agent can execute the action $shoot$, but the opponent does not, so the opponent is able to win the game from any state of any instance if he knows the right strategy. The agent and the adversary share the set of actions ~\{$up$, $right$, $down$, $left\}$. To reach the goal state, the agent has to be in the same row or column as the adverary and execute the action $shoot$. When the agent executes the $shoot$ action on the adversary's line, the adversary disappears. We can define the goal state, as the adversary not being on the board. The agent have one actions per turn and the adversary two, so that the agent can stand in the opponent's row or column and shoot before the opponent escapes.

@@

\subsubsection{Policy}
\begin{Verbatim}[fontsize=\footnotesize]
  1. A>0 AND B=0 AND C=0 AND D>0 -> {B ·, C ·, D ↓}
  2. A>0 AND B=0 AND C>0 AND D>0 -> {B ↑, C ↓, D ·}
  3. A>0 AND D=0 -> {A ↓}
  4. A>0 AND B>0 -> {A ↓}
\end{Verbatim}

\subsubsection{Policy Description}
According to the fourth rule of the policy, if the agent is in the same column as the agent it is good to decrease $A$, i.e. shoot the adversary. In the same way, according to the third rule, if the agent is in the same row as the adversary it is good to execute the action $shoot$. According to the second rule, if the agent is in the row adjacent to the adversary it is good to get in the same row. Finally, according to the first rule, if the agent is not in the row adjacent to the adversary, and is not in the same row or column, it is good to decrease the horizontal distance between the agent and the adversary. The first rule can also be read as, it is good for the agent to approach the adversary horizontally if you do not have the possibility to stand, or you are already in a position where it is good to shoot.

\subsubsection{Example of Game}
\begin{Verbatim}[fontsize=\footnotesize]
  A....... # .A...... # ..A..... # ..A..... # ...A.... # ....A... # ....A... # .....A.. # ......A. #
  ........ # ........ # ........ # ........ # ........ # ........ # ........ # ........ # ........ #
  ........ # ........ # ........ # ........ # ........ # ........ # .....T.. # ......T. # ......T. #
  ........ # ........ # ........ # .....T.. # .....T.. # .....T.. # ........ # ........ # ........ #
  .....T.. # .....T.. # .....T.. # ........ # ........ # ........ # ........ # ........ # ........ # ...

@@

\subsubsection{Example of Game}
\begin{Verbatim}[fontsize=\footnotesize]
   # .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
   # .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
   # .  .  .  .  .  .  .  K  .  .  #  .  .  .  .  .  .  .  K  .  .  #
   # .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
   # .  .  .  .  .  .  .  .  .  k  #  .  .  .  .  .  .  .  .  Q  k  #
   # .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
   # .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
   # .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
   # .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
   # .  .  Q  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
   # .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
   # .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
\end{Verbatim}

\subsection{Checkmate with Rook}
\subsubsection{Description}
In this game the action space is equal to the number of cells in the grid. In order to win the white pieces has to do a checkmate in one movement. In white pieces have the white-king and the white-rook, and black pieces have the black-king.

@@

>>>>

\subsection{Space Invaders}
\subsubsection{Description}
This domain is a simplified version of the original Space Invaders in which the Martians cannot shoot and cannot move below the penultimate row. In this domain the agent can $shoot$, move to the $right$ or to the $left$. The agent during his turn can do two actions, while the Martians can do one. Note that if the Martians and the agent have only one action per turn the agent cannot win from all possible states in the game, this is because of the situation where the agent stands in the Martian's column to shoot him on his next turn, but the next turn (the Martian's turn) the Martian can move away, this leads to an infinite loop if both are playing perfectly. That is, the agent must be able to move to the agent's column and then shoot. To win the agent has to kill all the martians, for this he must shoot each of them. The agent kills an opponent if he shoots while being in the same column, and he can only kill one for each shoot. The martians has three possible actions $left$, $right$, and $down$. All the adversaries move in the same direction. The agent cannot be destroyed in any way by the Martians, they can go down to the row above the agent. The agent cannot collide with the Martians as they cannot descend to the row above it. Note that the policy found is optimal and will solve all possible instances of the game. The policy is also optimal for a version of the game where the Martians destroy the agent when they reach the row above him.

@@

\subsubsection{Policy Description}
According to the first rule if the agent is not in the column of a Martian, $B$ is greater than zero, it is good to decrement $B$, i.e. move agent towards the closest column with martians. The second rule is applicable when the agent is under a Martian, in the same column, and according to this rule it is good to decrease the number of Martians, i.e. execute the action $shoot$.

@@

\subsection{Nim}
\subsubsection{Description}
In this domain there are two stacks (columns) of tokens. The number of actions available in each state depends on the number of tokens, a valid action involves removing one of the tokens. When a token is removed, that token and all the tokens above it are removed as well. When a token is removed, the turn is changed and the other player must remove a token as an action if there is one. The objective of the game is to leave the opponent with no actions available on his turn, i.e., without any token. The agent and the adversary has on action per turn.
It is interesting to note that in this game not all states are solvable if the opponent plays optimally. Knowing that the expert policy is to leave the opponent with both stacks with the same number of tokens, this state with both stacks leveled will not be solvable. Therefore, in order to solve it, we have made use of the $Bad(state)$ feature, which allows us not to have to define a good transition for each state, and as an alternative to label it as a bad state. The advantage is that if the initial state is solvable, bad states can always be avoided. However the policy is not complete for all states, in particular there will be no good transition for states with the two token stacks leveled, as we will see below, which we can say is the optimal representation of the game strategy. The winning tactic in this game is to level the two stacks of tokens by removing the token next to an empty cell and diagonally above another token, como se puede ver en~\cite{silver2020few}. The intuition behind this tactic is that if we follow it we can always reach a final situation where the opponent is forced to remove all the tokens from one pile, then we can remove the remaining tokens from the other pile.

@@

\subsubsection{Features}
\begin{Verbatim}[fontsize=\footnotesize]
  (A) Num[And(cell-has-token,Exists(left_cell,cell-has-token))] [k=5]
  (B) Num[And(cell-has-token,Exists(leftdown_cell,cell-has-token))] [k=5]
  (C) Num[And(cell-has-token,Exists(rightdown_cell,cell-has-token))] [k=5]
\end{Verbatim}

@@

\subsubsection{Policy}
\begin{Verbatim}[fontsize=\footnotesize]
  1. B>0 AND C>0 -> {A ·, B ↓, C ·}, {A ·, B ·, C ↓}
\end{Verbatim}

\subsubsection{Policy Description}
The policy guarantees to leave both stacks with the same number of tokens. This is done by decreasing the number of tokens that have a token to their leftdown or rightdown without decreasing the number of tokens with a token next to them.
As the policy clearly shows, there is no good transition, that includes a change in the number of tokens that a token has on its left. In other words, states where the same number of tokens in the two stacks cannot be preserved with some legal move are bad, and therefore they have no good transition. In fact according to the policy it is good to leave the opponent in these states bad, because there will be no legal move for the opponent to leave us with the same number of tokens in the two stacks.

@@

In this domain the agent has to pick up a missed pet and take it to a certain location, the destination. The pet has the option on its turn to escape from its captor to any of the cells adjacent to the pet, even if is holded by the agent it can escape. The goal is defined by the pet being droped at the destination cell, not holded by the agent. In this domain we use the additional predicate $holding_pet$ to indicate that the agent is holding the pet. The agent has two actions per turn and the adversary one, otherwise the game would not be solvable for all the states. The set of possile actions for the agent is ~\{$pick$, $drop$, $rightup$, $reightdown$, $leftup$, $leftdown\}$. The agent, in order to pick up the pet has to execute the action $pick$ adjacent to the the pet's cell. The agent, in order to drop the pet has to execute the action $drop$ while $holding_pet$, and the pet will be droped at de agent's cell. The set of possile actions for the pet consist in escape to one of the adjacent cells to it, i.e. at most eight possible actions depending on the position of the pet in the grid. The pet can not move into the delivery point.

@@

\subsubsection{Policy Description}
According to the first rule, it is good to move towards the pet if the agent is not holding it and the distance to it is greater than zero. According to the second rule, if the agent has the pet, and is at the destination, it is good to move towards the destination. According to the third rule of the policy, if the agent is holding the pet, and is at destination it is good to drop the pet. Finally, according to the fourth rule of the policy, if the agent does not have the pet, and is at the adjacent cell it is good pick it.

\subsubsection{Example of Game}
\begin{Verbatim}[fontsize=\footnotesize]
  p.... # p.... # ..... # ..... # ..... # ..... # ..... # ..... # ..... # ..... # ..... # ..... #
  ..... # .A... # .A... # ..... # ..... # ..... # ..... # ..... # ..... # ..... # ..... # ..... #
  A.... # ..... # ..... # ..A.. # .pA.. # ..A.. # ..... # ..... # ..... # ..... # ..... # ..... #
  ..... # ..... # ..... # ..... # ..... # ..... # ...A. # ..pA. # ...A. # ..... # ..... # ..... #
  ....D # ....D # ....D # ....D # ....D # ....D # ....D # ....D # ....D # ...pA # ....A # ....A #
\end{Verbatim}

\subsection{Shoot}
\subsubsection{Description}
In this domain an agent has to shoot an adversary. The goal is defined by the agent being alone on the grid, having finished with the adversary. The agent can execute the action $shoot$, but the opponent does not, so the opponent is able to win the game from any state of any instance if he knows the right strategy. The agent and the adversary share the set of actions ~\{$up$, $right$, $down$, $left\}$. To reach the goal state, the agent has to be in the same row or column as the adverary and execute the action $shoot$. When the agent executes the $shoot$ action on the adversary's line, the adversary disappears. We can define the goal state, as the adversary not being on the board. The agent have one actions per turn and the adversary two, so that the agent can stand in the opponent's row or column and shoot before the opponent escapes.

@@

\subsubsection{Policy}
\begin{Verbatim}[fontsize=\footnotesize]
  1. A>0 AND B=0 AND C=0 AND D>0 -> {A ·, B ·, C ·, D ↓}
  2. A>0 AND B=0 AND C>0 AND D>0 -> {A ·, B ↑, C ↓, D ·}
  3. A>0 AND D=0 -> {A ↓}
  4. A>0 AND B>0 -> {A ↓}
\end{Verbatim}

\subsubsection{Policy Description}
According to the fourth rule of the policy, if the agent is in the same column as the agent it is good to decrease $A$, i.e. shoot the adversary. In the same way, according to the third rule, if the agent is in the same row as the adversary it is good to execute the action $shoot$. According to the second rule, if the agent is in the row adjacent to the adversary it is good to get in the same row. Finally, according to the first rule, if the agent is not in the row adjacent to the adversary, and is not in the same row or column, it is good to decrease the horizontal distance between the agent and the adversary. The first rule can also be read as, it is good for the agent to approach the adversary horizontally.

\subsubsection{Example of Game}
\begin{Verbatim}[fontsize=\footnotesize]
  # A....... # .A...... # ..A..... # ..A..... # ...A.... # ....A... # ....A... # .....A.. # ......A. #
  # ........ # ........ # ........ # ........ # ........ # ........ # .....T.. # ......T. # ......T. #
  # ........ # ........ # ........ # ........ # ........ # ........ # ........ # ........ # ........ #
  # .....T.. # .....T.. # .....T.. # ........ # ........ # ........ # ........ # ........ # ........ #
  # ........ # ........ # ........ # .....T.. # .....T.. # .....T.. # ........ # ........ # ........ # ...

@@

\subsubsection{Task instance distribution}
To generate all the possible states from which the queen can checkmate the black king in one, in a board of $NxM$, we follow the following procedure. We generate all the possible checkmates that can be given to the black king if it is in the first row, then we apply three rotations to each board, 90, 180 and 270 degrees. To generate the checkmate with the black king in the first row we must take into account that for each black king position in this row there are three possible positions of the white king that allow a checkmate in one. In all these possible positions of the two kings we place the king in all the squares where it can checkmate the black king, and eliminate those where it would be giving a check. For the training, one hundred random boards of five by fifty are selected. To check that the policy is complete, it is tested on all possible, for each board size $NxM$ where both $N$ and $M$ are numbers no greater than twenty randomly selected.

\subsubsection{Example of Game}
\begin{Verbatim}[fontsize=\footnotesize]
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  K  .  .  #  .  .  .  .  .  .  .  K  .  .  #
# .  .  .  .  .  .  .  .  .  k  #  .  .  .  .  .  .  .  .  Q  k  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  Q  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
\end{Verbatim}

\subsection{Checkmate with Rook}
\subsubsection{Description}
In this game the action space is equal to the number of cells in the board. In order to win the white pieces has to do a checkmate in one movement. In white pieces have the white-king and the white-rook, and black pieces have the black-king.

@@

\subsubsection{Task instance distribution}
To generate all the possible states from which the rook can checkmate the black king in one, in a board of $NxM$, we follow the following procedure. We generate all the possible checkmates that can be given to the black king if it is in the first row, then we apply three rotations to each board, 90, 180 and 270 degrees. To generate the checkmate with the black king in the first row we must take into account that for each black king position in this row there are one possible position of the white king that allow a checkmate in one. In all these possible positions of the two kings we place the king in all the squares where it can checkmate the black king, and eliminate those where it would be giving a check. For the training, one hundred random boards of five by fifty are selected. To check that the policy is complete, it is tested on all possible, for each board size $NxM$ where both $N$ and $M$ are numbers no greater than twenty randomly selected.
